---
title: "ESM 206 Lab5"
author: "Ruoyu Wang"
date: "2019/10/29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

pulling frequently
commit save remotely but not on Git,but will help you find conflict

### Lab 5 Objectives

- learn to parse dates with 'lubridate'
- get counts of observations with count()
- uncount() rows
- one and two-sample t-text
- Create a hear map with geom_tile()
```{r, include=FALSE}
# read in data and attach packages
library(tidyverse)
library(janitor)
library(here)

# clean up data
lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  janitor::clean_names()
  
```

Use tidyr::uncount() function to convert our lobster datafrom frequency format to case format

not every observations in each row
if there are zero count, it will be removed by uncount
complete case function?  will contain zero
```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
```


#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

# defult theoetical value follows the normal distribution, only linear relationship matters
ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)
```

#### convert the 'date' column to class 'Data'

We'll use the 'lubridate' package to convert to Date format, and then to help us earily parse month & year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```

Now let's parse year and month using lubridate::month() and lubridate:: year()

```{r}
# month() will know the order even they are not shown in number. is better than month.abb which nned to manually add the order
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
```

Now let's find counts of obseved lobsters based on different hierarchical groupings:

First, lets count lobsters by each year and month

```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_ym

lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_y

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)

lobster_site
```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by + n()

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    lobster_number = n(),
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE)
  )

lobster_summary
```

#### Find confidence intervals

Use the t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples (...this will be vonered conceptually in lectures week 6)

```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)
# true population mean ?
```

#### Two-sample t-test to compare means

if a draw two samples from a same population with the same mean, their difference of means should be zero if the population is perfect

We want to ask: if there a significant difference in lobster lengths at Naples and Mohwk reefs?

We've done our necessary exploratory analyses to ditermine that a 2=sample t-test for means comparison is appropriate.

```{r}
napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)
mn_ttest
# true difference in means -- confidence interval(mean1-mean2)
```

THere is a sighificant difference in lobster lengths between Naples and Mohwk Reef.(t(df = statistic, p < 0.001, alpha = 0.05))

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK")) 

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2

mn_ttest2$conf.int
mn_ttest2$p.value
mn_ttest2$statistic
```

There is a sig diff (t(1850.8) = -19.849, p = 2.2e-16). bad for producibility

Here is the p-value: `r mn_ttest2$p.value`

here is my t-statisitc: `r mn_ttest2$statistic`

#### Now: a heatmap!

```{r}
lobster_ys <-  lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site))+
  geom_tile(aes(fill = n))
```

